---
title: "Data Summary"
author: "Seth Zissette"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
#' Code to create a data summary report
#' 
#' @param dat a \code{data.table}.
#' @param grace_pd_wks a \code{numeric} indicating the length of the grace period

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#' Load data
# dat

#' Set length of grace period
grace_pd_wks <- 4

# Data prep
dat[, enrollment_year := format(as.Date(dat[,enroll_date]), "%Y")]
dat[, wks_enrollment_to_art := as.numeric((art_start_date - enroll_date) / 7)]
dat[, wks_enrollment_to_tpt := as.numeric((tpt_start_date - enroll_date) / 7)]
dat[, wks_enrollment_to_tb := as.numeric((tb_diagnosis_date - enroll_date) / 7)]
dat[, tpt_in_grace_pd_vs_else := factor(ifelse(!is.na(wks_enrollment_to_tpt) & wks_enrollment_to_tpt <= grace_pd_wks, "TPT in grace period",
                                               "No TPT ever or TPT outside of grace period"),
                                        levels = c("TPT in grace period", "No TPT ever or TPT outside of grace period"),
                                        labels = c("TPT in\ngrace\nperiod\n", "No TPT\never\nor TPT\noutside\nof grace\nperiod\n"))]
dat[, tpt_in_grace_pd_vs_out_vs_never := factor(ifelse(!is.na(wks_enrollment_to_tpt) & wks_enrollment_to_tpt <= grace_pd_wks, "TPT in grace period",
                                                       ifelse(!is.na(wks_enrollment_to_tpt) & wks_enrollment_to_tpt > grace_pd_wks, "TPT outside of grace period",
                                                              "No TPT")),
                                                levels = c("TPT in grace period", "TPT outside of grace period", "No TPT"),
                                                labels = c("TPT in\ngrace\nperiod\n", "TPT\noutside\nof grace\nperiod\n", "No TPT\n"))]

year_min <- min(dat[,enrollment_year])
year_max <- max(dat[,enrollment_year])
```

## Tabular Data Summary

```{r tables}
tbl_summary(dat,
            missing = "no") %>%
  add_n(statistic = "{n_miss} ({p_miss}%)", last=T) %>%
  modify_header(label = "**Variable**",
                n = "**Missing**")
```

\newpage

## Time from Enrollment to ART Start by Year
Years are omitted if they contained no records who initiated ART.

```{r figures_art}
hist(dat[,wks_enrollment_to_art],
     main = "Enrollment to ART start - All Years", xlab = "Weeks",
     col = rgb(red = 0, green = 0, blue = 1, alpha = 0.5),
     border = rgb(red = 0, green = 0, blue = 1, alpha = 0.5))
for(i in year_min:year_max){
  if(all(is.na(dat[enrollment_year == i, wks_enrollment_to_art])) == FALSE){
    hist(dat[enrollment_year == i, wks_enrollment_to_art],
         main = paste0("Enrollment to ART start - ",i), xlab = "Weeks",
         col = rgb(red = 0, green = 0, blue = 1, alpha = 0.5),
         border = rgb(red = 0, green = 0, blue = 1, alpha = 0.5))
  }
}
```

\newpage

## Time from Enrollment to TPT Start by Year
Years are omitted if they contained no records who initiated TPT.

```{r figures_tpt}
hist(dat[,wks_enrollment_to_tpt],
     main = "Enrollment to TPT start - All Years", xlab = "Weeks",
     col = rgb(red = 0, green = 1, blue = 0, alpha = 0.5),
     border = rgb(red = 0, green = 1, blue = 0, alpha = 0.5))
for(i in year_min:year_max){
  if(all(is.na(dat[enrollment_year == i, wks_enrollment_to_tpt])) == FALSE){
    hist(dat[enrollment_year == i, wks_enrollment_to_tpt],
         main = paste0("Enrollment to TPT start - ",i), xlab = "Weeks",
         col = rgb(red = 0, green = 1, blue = 0, alpha = 0.5),
         border = rgb(red = 0, green = 1, blue = 0, alpha = 0.5))
  }
}
```

\newpage

## Comparison of Time from Enrollment to ART Start and TPT Start
Years are omitted if they contained no records who initiated TPT.

```{r figures_art_tpt}
hist(dat[,wks_enrollment_to_tpt],
     main = "Enrollment to ART and TPT start  - All Years", xlab = "Weeks",
     col = rgb(red = 0, green = 1, blue = 0, alpha = 0.5),
     border = rgb(red = 0, green = 1, blue = 0, alpha = 0.5))
hist(dat[,wks_enrollment_to_art],
     col = rgb(red = 0, green = 0, blue = 1, alpha = 0.5),
     border = rgb(red = 0, green = 0, blue = 1, alpha = 0.5),
     add = TRUE)
legend("topright", legend = c("ART start", "TPT start"),
       col = c(rgb(red = 0, green = 0, blue = 1, alpha = 0.5),
               rgb(red = 0, green = 1, blue = 0, alpha = 0.5)),
       pt.cex = 2, pch = 15)
for(i in year_min:year_max){
  if(all(is.na(dat[enrollment_year == i, wks_enrollment_to_tpt])) == FALSE){
    hist(dat[enrollment_year == i, wks_enrollment_to_tpt],
         main = paste0("Enrollment to ART and TPT start - ", i), xlab = "Weeks",
         col = rgb(red = 0, green = 1, blue = 0, alpha = 0.5),
         border = rgb(red = 0, green = 1, blue = 0, alpha = 0.5))
    hist(dat[enrollment_year == i, wks_enrollment_to_art],
         col = rgb(red = 0, green = 0, blue = 1, alpha = 0.5),
         border = rgb(red = 0, green = 0, blue = 1, alpha = 0.5),
         add = TRUE)
    legend("topright", legend = c("ART start", "TPT start"),
           col = c(rgb(red = 0, green = 0, blue = 1, alpha = 0.5),
                   rgb(red = 0, green = 1, blue = 0, alpha = 0.5)),
           pt.cex = 2, pch = 15)
  }
}
```

\newpage

## Time from Enrollment to TB Diagnosis by Year
Years are omitted if they contained no records who were diagnosed with TB.

```{r figures_tb}
hist(dat[,wks_enrollment_to_tb],
     main = "Enrollment date to TB diagnosis - All Years", xlab = "Weeks",
     col = rgb(red = 1, green = 0, blue = 0, alpha = 0.5),
     border = rgb(red = 1, green = 0, blue = 0, alpha = 0.5))
for(i in year_min:year_max){
  if(all(is.na(dat[enrollment_year == i, wks_enrollment_to_tb])) == FALSE){
    hist(dat[enrollment_year == i, wks_enrollment_to_tb],
         main = paste0("Enrollment date to TB diagnosis - ",i), xlab = "Weeks",
         col = rgb(red = 1, green = 0, blue = 0, alpha = 0.5),
         border = rgb(red = 1, green = 0, blue = 0, alpha = 0.5))
  }
}
```

\newpage

## Table 1 Summaries

### Comparing those who initiate TPT in grace period vs. everyone else
```{r tbl1_1}
tbl_summary(dat,
            by = tpt_in_grace_pd_vs_else,
            missing = "always",
            missing_text = "-Missing-") %>%
  add_overall() %>%
  modify_header(label = "**Variable***")
```

### Comparing those who initiate TPT in grace period vs. outside of grace period vs. never
```{r tbl1_2}
tbl_summary(dat,
            by = tpt_in_grace_pd_vs_out_vs_never,
            missing = "always",
            missing_text = "-Missing-") %>%
  add_overall() %>%
  modify_header(label = "**Variable***")
```
