---
title: Kenya PROTECT analysis
output: html_document
params:
	exclude_weeks: 4
	grace_pd_wks: 12
	admin_cens_wk: 104
	n_boot: 1000
	run_bootstrap: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      error = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(fastverse)
library(tidyverse)
library(future.apply)
library(progressr)
set.seed(404)

ncores <- parallel::detectCores()
future::plan('multisession', workers = ncores-1)
here::i_am("reports/kenya_analysis.Rmd")
```

# Loading data

First, we read in the weekly records data that was created previously.

```{r data-load}
weekly_records_data <- fread(here::here("data/weekly_records_data.csv"))
```

Next, we exclude early cases of TB. Here, we are using the `exclude_weeks` variable from `params` specified in the header. This value can be changed in the header of the R Markdown document to run a sensitivity analysis with a different window for TB case exclusion. The default value is 4 weeks.

```{r exclude-early-tb}
early_tb <- weekly_records_data$tb_wk != 99999 & weekly_records_data$tb_wk <= params$exclude_week
weekly_records_data_exclude <- weekly_records_data[!early_tb,]
```

Next, we remove rows from the weekly records data that are after the end of our defined follow-up period. This value can also be changed in the header of the R Markdown document and defaults to 104 weeks (i.e., 2 years).

```{r remove-weeks-after-admin-censoring}
weekly_records_data_exclude <- weekly_records_data_exclude[
  wk <= params$admin_cens_wk, 
]
weekly_records_data_exclude$admin_cens_wk <- params$admin_cens_wk
```

Next, we need to merge in the enrollment year to the weekly records data.

```{r mege-enroll-year}
enroll_year <- readRDS(
  here::here("../enroll_year_only.rds")
)
enroll_year <- as.data.table(enroll_year)
weekly_records_data_exclude <- merge(weekly_records_data_exclude, enroll_year, by = "id")
```

After this step, we have a data set that is ready for analysis.

# Fitting propensity models

The code chunk below defines the set of baseline covariates that are used in the analysis, as well as the formulas used to estimate the propensity score models.

```{r define-model-formulas}
baseline_covariates <- c(
  "martial_status", "gender", "age", "religion", "education",
  "known_tb_prior_to_enroll", "known_tpt_prior_to_enroll",
  "who_stage", "year"
)
cat(
	"Baseline covariates used in analysis: ", 
	paste0(baseline_covariates, sep = ", ")
)

denom_model_formula <- paste0(
  paste0(baseline_covariates, collapse = " + "), " + ",
  "tb_symptoms_measured_visit_k + I(tb_symptoms_measured_visit_k*tb_symptoms_visit_k)", "+",
  "art_init + I(art_init*art_adherence_visit_k)", "+",
  "have_cd4 + I(have_cd4*cd4_count)", "+",
  "I((wk < 8)*(year<2020)*number_encounters_last_8_weeks)", "+",
  "splines::ns(wk, df = 3)"
)
cat(
	"Model formula used for denominator propensity model: ",
	denom_model_formula
)

num_model_formula <- paste0(
  paste0(baseline_covariates, collapse = " + "), " + ",
  "tb_symptoms_measured_visit_k + I(tb_symptoms_measured_visit_k*tb_symptoms_visit_k)", "+",
  "art_init + I(art_init*art_adherence_visit_k)", "+",
  "have_cd4 + I(have_cd4*cd4_count)", "+",
  "I((wk < 8)*(year<2020)*number_encounters_last_8_weeks)", "+",
  "wk + I(wk^2)"
)
cat(
	"Model formula used for numerator propensity model: ",
	num_model_formula
)

right_cens_model_formula <- paste0(
  "splines::ns(wk, df = 3) + ",
  paste0(baseline_covariates, collapse = " + "), " + ",
  "tpt_init + ",
  "art_init + I(art_init*art_adherence_visit_k)"
)
cat(
	"Model formula used for right censoring model: ",
	right_cens_model_formula
)
```

Now we are ready to fit the propensity score models.

```{r fit-propensity-models}
source(here::here("code/fit_propensity_models.R"))

propensity_output <- fit_propensity_models(
  weekly_records_data = weekly_records_data_exclude,
  grace_pd_wks = params$grace_pd_wks,
  denom_model_formula = denom_model_formula,
  num_model_formula = num_model_formula,
  right_cens_model_formula = right_cens_model_formula
)
```

# Creating cloned data

Next we created the "cloned" data set.

```{r, create-cloned-data}
source(here::here("code/create_cloned_data_set.R"))

cloned_data_sets <- create_cloned_data_set(
  propensity_output
)
```

# Fitting MSMs

In the next code chunk, we define the model formulas that are used to fit the MSMs.

```{r, define-msm-formulas}
msm_formula <- "z + splines::ns(wk, knots = c(4, 26, 52, 78))"
```

Now we fit MSMs for the endpoints of interest.

```{r, fit-tb-msm}
source(here::here("code/fit_msm.R"))

msm_fit_tb <- fit_msm(
  cloned_data_set = cloned_data_sets$tb,
  msm_formula = msm_formula,
  return_msm_model = TRUE,
  return_msm_vcov = FALSE,
  gee = FALSE
)

cat(
	"Coefficients for the TB model: ",
	msm_fit_tb$msm_coef
)

msm_fit_death <- fit_msm(
  cloned_data_set = cloned_data_sets$death,
  msm_formula = msm_formula,
  return_msm_model = TRUE,
  return_msm_vcov = TRUE,
  gee = TRUE  
)

cat(
	"Coefficients for the death model: ",
	msm_fit_death$msm_coef
)

msm_fit_death_for_tb <- fit_msm(
  cloned_data_set = cloned_data_sets$death_for_tb,
  msm_formula = msm_formula,
  return_msm_model = TRUE
)

```


# Fitting MSMs to assess effect heterogeneity

It is also of interest to evaluate the possible presence of effect heterogeneity. Below we define a model formula used to assess effect heterogeneity with WHO stage at enrollment.

```{r, msm-formula-for-effect-hetero}
msm_formula_hetero <- "splines::ns(wk, knots = c(4, 26, 52, 78)) + z*factor(who_stage)"
```

```{r, msms-for-effect-hetero}
msm_fit_tb_hetero <- fit_msm(
	cloned_data_set = cloned_data_sets$tb, 
	msm_formula = msm_formula_hetero,
	return_msm_model = TRUE,
	gee = FALSE,
	return_msm_vcov = FALSE
)

cat(
	"Coefficients for the TB effect heterogeneity model: ",
	msm_fit_tb_hetero$msm_coef
)

msm_fit_death_hetero <- fit_msm(
	cloned_data_set = cloned_data_sets$death, 
	msm_formula = msm_formula_hetero,
	return_msm_model = TRUE,
	gee = FALSE,
	return_msm_vcov = FALSE
)

cat(
	"Coefficients for the death effect heterogeneity model: ",
	msm_fit_death_hetero$msm_coef
)

msm_fit_death_for_tb_hetero <- fit_msm(
	cloned_data_set = cloned_data_sets$death_for_tb, 
	msm_formula = msm_formula_hetero,
	return_msm_model = TRUE,
	gee = FALSE,
	return_msm_vcov = FALSE
)
```

# Summarizing outputs of MSMs

The following code conversts the MSM output into estimates of cumulative incidence.

```{r summarize-cum-inc}
source(here::here("code/primary_analysis/compute_cuminc.R"))

cuminc <- compute_cuminc(
	msm_fit_tb = msm_fit_tb,
	msm_fit_death = msm_fit_death,
	msm_fit_death_for_tb = msm_fit_death_for_tb,
	max_wk = params$admin_cens_wk
)

cuminc_hetero <- compute_cuminc(
	msm_fit_tb = msm_fit_tb_hetero,
	msm_fit_death = msm_fit_death_hetero,
	msm_fit_death_for_tb = msm_fit_death_for_tb_hetero,
	max_wk = params$admin_cens_wk
)
```

# Bootstrapping

Now we are ready to run the bootstrap code to get confidence intervals for all results. This code will take quite some time to run

```{r fit-bootstrap}
source(here::here("code/primary_analysis/bootstrap.R"))

bootstrap_results <- run_bootstrap(
	nboot = params$n_boot,
	# ensure all these inputs are same as above!
  weekly_records_data = weekly_records_data_exclude,
  grace_pd_wks = params$grace_pd_wks,
  denom_model_formula = denom_model_formula,
  num_model_formula = num_model_formula,
  right_cens_model_formula = right_cens_model_formula,
  admin_cens_wks = params$admin_cens_wk, 
  # input all msm formulas that are of interest
  msm_formulas_tb = c(msm_formula, msm_formula_hetero),
  msm_formulas_death = c(msm_formula, msm_formula_hetero),
  msm_formulas_death_for_tb = c(msm_formula, msm_formula_hetero)
)
```

After the bootstrap process has completed, we save the results to ensure results are not lost. E.g.,

```{r save-bootstrap}
saveRDS(
	bootstrap_results,
	here::here("data/bootstrap_results.rds")
)
```

Finally, confidence intervals and other relevant information for inference can be obtained using the `get_bootstrap_ci` function.

```{r bootstrap_ci}
bootstrap_ci <- get_bootstrap_ci(bootstrap_results)
```

# Saving all results

The final set of outputs can be combined into a `list` named `results` that combines the `propensity_output`, `cf_init_dist`, `msm_fit`, `cuminc`, `bootstrap_results`, and `bootstrap_ci` objects. The code below tries to reduce the size of these objects as much as possible prior to saving.

```{r combine_results}
# Null out parts of propensity_output no longer needed
propensity_output$weekly_records_data <- NULL
propensity_output[["models"]][["denom_model"]]$qr <- NULL
propensity_output[["models"]][["num_model"]]$qr <- NULL
propensity_output[["models"]][["cens_model_tb"]]$qr <- NULL
propensity_output[["models"]][["cens_model_death"]]$qr <- NULL

# Combine MSMs and null out parts no longer needed
models <- mget(ls(pattern = paste0("^", "msm")))
for(i in seq_len(length(models))){
  models[[i]][["msm_model"]]$qr <- NULL
}

# Combine cumulative incidences
cumincs <- mget(ls(pattern = paste0("^", "cuminc")))

# Combine all results
results <- mget(c(ls(pattern = paste0("^", "propensity")),
                  ls(pattern = paste0("^", "cf_init")),
                  ls(pattern = paste0("^", "models")),
                  ls(pattern = paste0("^", "cumincs")),
                  ls(pattern = paste0("^", "bootstrap_r")),
                  ls(pattern = paste0("^", "bootstrap_ci"))))
```

The results can be saved in `.rds` format, as shown below.

```{r save_results}
saveRDS(
	results,
	here::here("data/R_results_2.rds")
)
```
